Topological Equivariant Artist Model for Visualization Library Architecture

### 2021-10-11 paper/main.tex
thursday meeting on stalks + notes
sheaves

We define the categories $\mathcal{\dtotal}, \mathcal{\gtotal}$ as sheaves of sections\cite{SheafMathematics2021,spanier1989algebraic}, of fiberbundles, as introduced in \autoref{sec:related-work:fiber-bundles}.  We assume that $\mathcal{E}$ is defined in terms of a fixed base and fiber space, for example a class of images or tables with the same schema. The graphic category $\mathcal{\gtotal}$ consists of the set of graphics that an artist \vartist is allowed to generate from a given $\mathcal{E}$, which is discussed further in \autoref{sec:artist:construction}.

@@

\begin{definition} The category $\mathcal{\dtotal}$ consists of 
  \begin{itemize}
    \item\textit{object} the contravariant functor $\sheaf: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
    \item \textit{morphisms} the endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $\bullet: \dtotal\rightarrow \dtotal$, which are the fiber maps $\monoid:\mathcal{\dfiber}\rightarrow \mathcal{\dfiber}$ when the sheaf is over a single point in the base space $\dbasepoint\in \dbase$
  \end{itemize}
\end{definition}

The functor object of $\mathcal{E}$ is a presheaf because the definition of a presheaf is that it is a contravariant functor from a category $\mathcal{K}$ to a set \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}. As illustrated in the diagram, 
\begin{equation}
  \label{eq:artist:[presheaf}
  \begin{tikzcd}[row sep=small]
    U_1 \arrow[dd, "\iota"'] & {\Gamma(U_1, \dtotal)}                              \\
    {} \arrow[r, "\sheaf"]   & {}                                                  \\
    U_2                      & {\Gamma(U_2, \dtotal)} \arrow[uu, "\iota^*"']
    \end{tikzcd}
\end{equation}
the functor \sheaf\ is contravariant because the inclusion map $\iota$ between objects in $\mathcal{K}$ goes the in the opposite direction from the map $\iota$ between open set objects in $\mathcal{K}$ \cite{riehlCategoryTheoryContext}. The open set $U_1$ is a subset of $U_2$, therefore $\iota: U_1 \rightarrow U_2$. On the other hand, a set of functions defined on a space $U_2$ must be defined on the subspace $\iota: \Gamma(U_2, \dbase) \rightarrow \Gamma(U_1, \dbase)$ but the reverse does not need to be true. For example, a continuous function defined over the interval $\left[.25,.5\right]$ need not be defined over $\left[0, 1\right]$ but must exist over  $\left[.3, .45\right]$. We use the notation $\mathcal{K}^{op}$ to signify that the functor $\sheaf$ is contravariant and $\sheaf$ to denote that it is a sheaf. A sheaf is a presheaf where sections are defined over unions of open sets \cite{nlab:sheaf,ghristElementaryAppliedTopology2014}

The graphic category $\mathcal{\gtotal}$ consists of the functor $\sheaf: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$. The category $\mathcal{S}$ is the category of open sets over the graphic base space \gbase\, and $\Gamma(\mathcal{S}, \gtotal)$ is the set of sections $\gsection$ over $\mathcal{\gbase}$. The artist is a morphism of sheafs $\vartist: \mathcal{\dtotal} \rightarrow \mathcal{\gtotal}$, which means it is by definition a natural transform\cite{SheafMathematics2021}. This means that the following diagram must commute

\begin{equation}
  \label{eq:artist:natural-transform}
  \begin{tikzcd}
  \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "\monoid"'] & \mathcal{\gtotal} \arrow[d, "\vartist(\monoid)"] \\
  \mathcal{\dtotal} \arrow[r, "\vartist"]                       & \mathcal{\gtotal}                               
  \end{tikzcd}
\end{equation}

such that the morphisms on the graphic category $\mathcal{H}$ are homomorphic to the morphisms $\monoid$ on the fibers $\mathcal{F}$ \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}. In \autoref{sec:artist:construction}, we show a method of constructing the artist such that this constraint can be satisfied, thereby ensuring equivariance. We then use this formulation to guide the development of visualization library components in \autoref{sec:case-study}.

@@

\begin{min

>>>>

We define the categories $\mathcal{\dtotal}$ and $\mathcal{\gtotal}$ as sheaves of sections\cite{SheafMathematics2021,spanier1989algebraic}, of fiberbundles, as introduced in \autoref{sec:related-work:fiber-bundles}. Sheaves are mathematical abstractions of how data over different parts of a continuous base space are glued together \cite{ghristElementaryAppliedTopology2014, ghri}, whether the data is distributed, streaming, or in disk; therefore sheaves are a way to express the continuity that the artist preserves. 
We assume that $\mathcal{E}$ is defined in terms of a fixed base and fiber space, for example a class of images or tables with the same schema. The graphic category $\mathcal{\gtotal}$ consists of the set of graphics that an artist \vartist\ is allowed to generate from a given $\mathcal{E}$, which is discussed further in \autoref{sec:artist:construction}.

@@

\begin{definition} The category $\mathcal{\dtotal}$ consists of 
  \begin{itemize}
    \item\textit{object} the sheaf\cite{SheafMathematics2021,ghristElementaryAppliedTopology2014} $\sheaf: \mathcal{K}^{op} \rightarrow \Gamma(\mathcal{K}, \dtotal)$
    \item \textit{morphisms} the endomorphic\cite{fongInvitationAppliedCategory2019} bundle maps $\bullet: \dtotal\rightarrow \dtotal$, which are the fiber maps $Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ when the sheaf is restricted to the stalk\cite{StalkSheaf2019} $\mathscr{F}_{k}$ over a point $\dbasepoint\in \dbase$
  \end{itemize}
\end{definition}

The functor object of $\mathcal{E}$ is a presheaf because the definition of a presheaf is that it is a contravariant functor from a category $\mathcal{K}$ to a set \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}. As illustrated in the diagram, 
\begin{equation}
  \label{eq:artist:[presheaf}
  \begin{tikzcd}[row sep=small]
    U_1 \arrow[dd, "\iota"'] & {\Gamma(U_1, \dtotal)}                              \\
    {} \arrow[r, "\sheaf"]   & {}                                                  \\
    U_2                      & {\Gamma(U_2, \dtotal)} \arrow[uu, "\iota^*"']
    \end{tikzcd}
\end{equation}
the functor \sheaf\ is contravariant because the inclusion map $\iota$ between open sets in $\mathcal{K}$ goes in the opposite direction from the map $\iota^*$ between the sections in $\Gamma(U_{i}, \dtotal)$ \cite{riehlCategoryTheoryContext}. The open set $U_1$ is a subset of $U_2$, therefore $\iota: U_1 \rightarrow U_2$. On the other hand, a set of functions defined on a space $U_2$ must be defined on the subspace $\iota: \Gamma(U_2, \dbase) \rightarrow \Gamma(U_1, \dbase)$ but the reverse does not need to be true. For example, a continuous function defined over the interval $\left[.25,.5\right]$ need not be defined over $\left[0, 1\right]$ but must exist over  $\left[.3, .45\right]$. We use the notation $\mathcal{K}^{op}$ to signify that the functor $\sheaf$ is contravariant and $\sheaf$ to denote that it is a sheaf. A sheaf is a presheaf where sections are defined over unions of open sets \cite{nlab:sheaf,ghristElementaryAppliedTopology2014}. The sheaf over a limit of open sets that contain a point \dbasepoint\ is approximately the same as a sheaf over a point \dbase\ and is called the stalk $\mathscr{F}_{k}\coloneqq \lim\limits_{U\ni k} \mathscr{F}(U)$ \cite{StalkSheaf2019}. The fiber space over a point is embedded in the stalk over that point $\dfiber_{k}\hookrightarrow \mathscr{F}_{k}$. Our criteria for equivariance is that the set of maps from fiber to fiber $Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ is preserved, since these maps include the types of data transformations described in \autoref{sec:related-work:equivariance}. When a fiber bundle is trivial, meaning $\dtotal=\dbase \times \dfiber$, then we can dispense with localization and directly consider the fiber maps $Hom(F_{k}, F_{k})$. In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.

The graphic category $\mathcal{\gtotal}$ consists of the functor $\sheaf: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$. The category $\mathcal{S}$ is the category of open sets over the graphic base space \gbase\, and $\Gamma(\mathcal{S}, \gtotal)$ is the set of sections $\gsection$ over $\mathcal{\gbase}$. The artist is a morphism of sheafs $\vartist: \mathcal{\dtotal} \rightarrow \mathcal{\gtotal}$, which means it is by definition a natural transform\cite{SheafMathematics2021}. \note{which means? why is to good for it to be a natural transformation->}

%% 
Extra property that implies equivariance is preserved is illustrated in the diagram which must commute
%% note Im_A in \H, A_M (lifting/pushforward)
\begin{equation}
  \label{eq:artist:natural-transform}
  \begin{tikzcd}
  \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "\monoid"'] & \mathcal{\gtotal} \arrow[d, "\vartist(\monoid)"] \\
  \mathcal{\dtotal} \arrow[r, "\vartist"]                       & \mathcal{\gtotal}                               
  \end{tikzcd}
\end{equation}

%H is the codomain, ImA is the range, K & S -> 
% clipping problem: shift lines and stuff w/o shifting frame, when we shift the frame, only shifting the bits of H that belongs to the artist  
such that the morphisms on the graphic category $\mathcal{H}$ are homomorphic to the morphisms $\monoid$ on the fibers $\mathcal{F}$ \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}. \note{equivariant not homomorphic, this isn't saying anything, change Hom(F,F)}

In \autoref{sec:artist:construction}, we show a method of constructing the artist such that this constraint can be satisfied, thereby ensuring equivariance. We then use this formulation to guide the development of visualization library components in \autoref{sec:case-study}.

@@

\begin{minted}{python}
\end{minted}
